Convex Surrogates via Polyhedral Losses

### 2019-02-02 cvx-fin.tex
sec 2 done

@@

We write loss functions as $L:\R\to\reals^\Y_+$ which map a report (prediction) $r\in\R$ to the vector of loss values $L(r) = (L(r)_y)_{y\in\Y}$ for each possible outcome $y\in\Y$.
Often $\R$ will be taken to be a finite set for the given discrete loss (in which case we will write $\ell$ not $L$), and $\reals^d$ for the surrogate.
The expected loss of report $r$ when $y \sim p$ is written $\inprod{p}{L(r)}$.
For example, 0-1 loss is a discrete loss with $\R = \Y$ given by $\ell(r)_y = \Ind[r \neq y]$.

@@

Thus, $\Gamma(p)$ is the set of reports which should be optimal for a given distribution $p$, and $\Gamma_r$ is the set of distributions for which the report $r$ should be optimal.
For example, the \emph{mode} is the property $\mode(p) = \argmax_{y\in\Y} p_y$, which is finite.
The mode captures the set of optimal reports for 0-1 loss: for each distribution over the labels, one should report the most likely label.
In this case we say 0-1 loss \emph{elicits} the mode, as we now formalize.

\begin{definition}[Elicits]
  A loss $\ell:\R\to\reals^\Y_+$, \emph{elicits} a property $\gamma:\simplex \toto \R$ if for all $p \in \simplex$,
  \begin{equation}
  \Gamma(p) = \argmin_{r \in \R} \inprod{p}{L(r)}~.
  \end{equation}
  \bo{add $\gamma[\ell]$ notation}
\end{definition}
This terminology comes from the information elicitation literature, in which a report $r$ is elicited from some forecaster by scoring her with a loss on the observed outcome $y$; she tries to minimize expected loss.

Typically in this paper, $\gamma$, $\R$, and $\ell$ will be used for a finite property, finite report space, and loss eliciting it, while $\Gamma$, $\R'$ and $L$ will be used for some other property, often with report space $\R' = \reals^d$.
In this case, we prefer $r$ for a member of $\R$ and $u$ for members of $\reals^d$.

In $\reals^d$, a \emph{polyhedral set} is the intersection of a finite number of halfspaces.
A \emph{polytope} is a bounded polyhedral set.
\begin{definition}[Polyhedral loss]
  A \emph{polyhedral function} is a pointwise maximum over a finite set of affine functions.
  In other words, it is a function whose epigraph (set of points lying above it) forms a polyhedral set.
  $L: \reals^d \to \reals^{\Y}_+$ is a \emph{polyhedral loss} if $L(u)_y$ is a polyhedral function of $u$ for each $y$.
\end{definition}

\subsection{Link Functions and Elicitation Complexity}\label{sec:link-elic}
Often, for computational or other reasons, one is interested in a property $\gamma$ such as the mode, but wishes to elicit it by minimizing some other ``surrogate'' loss $L$ over some other space such as $\reals^d$, then mapping the result back to $\R$ using a link function.
We now formalize this procedure.
\bo{Maybe mention that previous formalizations are for single-valued properties only if true? And if so, that they don't really apply to finite properties??}

\bo{Didn't touch these}
\begin{definition}\label{def:links}
  Let properties $\gamma:\simplex\toto\R$ and $\Gamma:\simplex\toto\R'$ be given.
  A \emph{link function} is a map $\psi:\R'\to\R$.
  We say that a link $\psi$ is:
  \begin{itemize}
  \item \emph{Calibrated} if for all $p\in\simplex$, $r'\in \Gamma(p) \implies \psi(r') \in \gamma(p)$; in other words, if for all $r'\in\R'$, $\Gamma_{r'} \subseteq \gamma_{\psi(r')}$;
  \item \emph{Separated}, with respect to a loss $L$ eliciting $\Gamma$, if for all $p \in \simplex$,
  \begin{align*}
  \inf_{u \in \R'; \psi(u) \not \in \gamma(p)} \E_{Y\sim p}[L(u, Y)] > \inf_{u \in \R'}\E_{Y\sim p}[L(u, Y)]~.
  \end{align*}
  \item \emph{Exhaustive} if for all $p\in\simplex$, $\gamma(p) = \psi(\Gamma(p)) := \{\psi(r') : r'\in\Gamma(p)\}$.
  \end{itemize}
\end{definition}

It is also interesting to contrast the definitions above with the following.
\begin{definition}
  A \emph{strong link} is a set-valued function $\psi:\R'\toto\R$ such that for all $p\in\simplex$ and $r'\in\R'$, we have $r'\in\Gamma(p) \implies \gamma(p) = \psi(r')$.
\end{definition}

\raf{We should note here that a discussion like this has not yet appeared in the property elicitation literature, since most papers address direct elicitation, where no link is needed, or single-valued properties in the case of indirect elicitation.  Cite examples for indirect: Frongillo--Kash (NIPS), Fissler--Ziegel (AoS), mode paper, others?}

\raf{Discussion about these definitions using 0-1 loss, hinge, and logistic.  Namely: sign is calibrated for hinge and logistic.  If sign is interpreted as set-valued (0 maps to $\{1,-1\}$, then it is strong for logistic.  If not, sign is not exhaustive for logistic.  Sign is exhaustive for hinge.  It is also separated, but something like $\psi(u) =
  \begin{cases}
    -1 & u \leq -1\\
    1 & u > -1
  \end{cases}$
  would not be, even though it would be calibrated.}

@@

\raf{Notation up for discussion}
$\eliccvx$ and $\elicpoly$; forward reference $\elicembed$

We define the elicitation complexities of various classes of loss functions as follows: elic$_\C(\gamma)$ is the minimum dimension $d$ such that $\gamma$ can be indirectly elicited by a loss function $L: \R'^d \to \reals^\Y_+$ in the class $\C$.
For example, if $\eliccvx(\gamma) \leq d$, then there is a \emph{convex} loss $L:\reals^d \to \reals^\Y_+$ that indirectly elicits $\gamma$.
Similarly, if $\elicpoly(\gamma) \leq d$, then there is a \emph{polyhedral} loss $L:\reals^d \to \reals^\Y_+$ that indirectly elicits $\gamma$.

\subsection{Important Examples}
  Ranking and classification problems are well-known to the field of machine learning.
  The most studied classification problem is the elicitation of the mode.
  While the mode over $n$ outcomes is elicitable by 0-1 loss, it is known that eliciting the mode by a convex loss is much more difficult.
  In fact, the best convex surrogate that can be used to elicit the mode ``embeds'' the reports into $\reals^{n-1}$.
  Therefore, we say $\eliccvx($mode$) = n-1$.

@@

\section{Polyhedral Losses and Embeddings}

@@

\begin{definition}
  We say $\gamma$ is \emph{$d$-embeddable}, or just \emph{embeddable}, if such an $L$ exists and is convex.
  The \emph{embedding dimension} of a finite property $\gamma$, written $\elicembed(\gamma)$, is the minimum $d\geq 0$ such that $\gamma$ is $d$-embeddable.
\end{definition}

@@

\section{Embeddings via Real-Valued Losses}

>>>>

\raf{MOVED FROM SEC 2; maybe incorporate?}
Often, for computational or other reasons, one is interested in a property $\gamma$ such as the mode, but wishes to elicit it by minimizing some other ``surrogate'' loss $L$ over some other space such as $\reals^d$, then mapping the result back to $\R$ using a link function.
We now formalize this procedure.

@@

We write loss functions as $L:\R\to\reals^\Y_+$ which map a report (prediction) $r\in\R$ to the vector of loss values $L(r) = (L(r)_y)_{y\in\Y}$ for each possible outcome $y\in\Y$.
Often $\R$ will be taken to be a finite set for the given discrete loss (in which case we will write $\ell$ not $L$), and $\reals^d$ for the surrogate.
The expected loss of report $r$ when $y \sim p$ is written $\inprod{p}{L(r)}$.
For example, 0-1 loss is a discrete loss with $\R = \Y = \{-1,1\}$ given by $\ell(r)_y = \Ind[r \neq y]$.
Two important examples of surrogates for 0-1 loss, with $\R=\reals$, are hinge loss $L(u)_y = \max(0,1-yu)$ and logistic loss $L(u)_y = \log(1+\exp(-yu))$.

@@

Thus, $\Gamma(p)$ is the set of reports which should be optimal for a given distribution $p$, and $\Gamma_r$ is the set of distributions for which the report $r$ should be optimal.
For example, the \emph{mode} is the property $\mode(p) = \argmax_{y\in\Y} p_y$, which is finite.
The mode captures the set of optimal reports for 0-1 loss: for each distribution over the labels, one should report the most likely label.
In this case we say 0-1 loss \emph{elicits} the mode, as we formalize below.
This terminology comes from the information elicitation literature~\citep{savage1971elicitation,osband1985information-eliciting,lambert2008eliciting}, in which a report $r$ is elicited from some forecaster by scoring her with a loss on the observed outcome $y$.

\begin{definition}[Elicits]
  A loss $L:\R\to\reals^\Y_+$, \emph{elicits} a property $\Gamma:\simplex \toto \R$ if
  \begin{equation}
    \forall p\in\simplex,\;\;\;\Gamma(p) = \argmin_{r \in \R} \inprod{p}{L(r)}~.
  \end{equation}
  As $\Gamma$ is uniquely defined by $L$, we write $\prop{L}$ to refer to the property elicited by a loss $L$.
\end{definition}

For finite properties and discrete losses, we will use lowercase notation $\gamma$ and $\ell$, respectively, while $\Gamma$ and $L$ will be used for general properties.
Often we will work with two properties simultaneously, a finite property $\gamma:\simplex\toto\R$, and its embedded version $\Gamma:\simplex\toto\reals^d$.
In this case, we write the embedded reports as $u\in\reals^d$, reserving $r\in\R$ for the finite property.

Most of the loss functions $L$ we consider will be piecewise linear and convex, or \emph{polyhedral}; we therefore will make use of the theory of convex polytopes.
In $\reals^d$, a \emph{polyhedral set} or \emph{polyhedron} is the intersection of a finite number of halfspaces.
A \emph{polytope} is a bounded polyhedral set.
A convex function $f:\reals^d\to\reals$ is \emph{polyhedral} if its epigraph is polyhedral, or equivalently, if it can be written as a pointwise maximum of a finite set of affine functions~\cite{rockafellar1997convex}.

\begin{definition}[Polyhedral loss]
  A loss $L: \reals^d \to \reals^{\Y}_+$ is a \emph{polyhedral} if $L(u)_y$ is a polyhedral (convex) function of $u$ for each $y\in\Y$.
\end{definition}
For example, note that hinge loss is polyhedral, as is the surrogate for abstain loss described in \S~\ref{sec:important-examples}.

\subsection{Link Functions and Elicitation Complexity}\label{sec:link-elic}

When designing a surrogate loss function, we require some sort of statistical consistency with the original loss, so that minimizing the surrogate is an effective alternative to directly minimizing the original loss.
We formalize this consistency below in terms of the properties elicited by each loss.
In particular, we seek a \emph{link} function, which maps optimal reports from the surrogate to the original reports, satisfying certain guarantees.
\bo{Maybe mention that previous formalizations are for single-valued properties only if true? And if so, that they don't really apply to finite properties??}

\begin{definition}[Calibrated Link]\label{def:links}
  Let properties $\gamma:\simplex\toto\R$ and $\Gamma:\simplex\toto\R'$ be given.
  A \emph{link function} is a map $\psi:\R'\to\R$.
  We say that a link $\psi$ is:
  \begin{itemize}
  \item \emph{Calibrated} if for all $p\in\simplex$, $r'\in \Gamma(p) \implies \psi(r') \in \gamma(p)$; in other words, if for all $r'\in\R'$, $\Gamma_{r'} \subseteq \gamma_{\psi(r')}$;
  \item \emph{Separated}, with respect to a loss $L$ eliciting $\Gamma$, if for all $p \in \simplex$,
  \begin{align*}
  \inf_{u \in \R'; \psi(u) \not \in \gamma(p)} \E_{Y\sim p}[L(u, Y)] > \inf_{u \in \R'}\E_{Y\sim p}[L(u, Y)]~.
  \end{align*}
  \end{itemize}
\end{definition}

Roughly speaking, a calibrated link requires that optimal surrogate reports map to optimal original reports, while a separated link adds the stronger guarantee that one cannot approach the optimal surrogate loss while making suboptimal original predictions.
A classic example of a link which is both calibrated and separated is $\psi = \sgn$ when $\gamma=\mode$ and $L$ is either hinge loss or logistic loss.
On the other hand, the link $\psi(u) =
\begin{cases}
  -1 & u \leq -1\\
  1 & u > -1
\end{cases}$
would not be separated for hinge loss, but it would still be calibrated, as the reports $u\in(-1,1)$ are optimal for hinge loss only when $p=(1/2,1/2)$, in which case both $-1$ and $1$ are optimal for the mode (0-1 loss).

\raft{Aside about other link definitions commented out}
% As a brief aside, it is interesting to contrast the definitions above with the following: a link is
% \emph{exhaustive} if for all $p\in\simplex$, $\gamma(p) = \psi(\Gamma(p)) := \{\psi(r') : r'\in\Gamma(p)\}$.
% A \emph{strong link} is a set-valued function $\psi:\R'\toto\R$ such that for all $p\in\simplex$ and $r'\in\R'$, we have $r'\in\Gamma(p) \implies \gamma(p) = \psi(r')$.
% If sign is interpreted as set-valued (0 maps to $\{1,-1\}$, then it is strong for logistic.  If not, sign is not exhaustive for logistic.  Sign is exhaustive for hinge.}

It is often useful to express the relationship between $L$ and $\gamma$ above directly, which is the notion of \emph{indirect elicitation}.

@@

Often one is given a property $\gamma$ (typically via a discrete loss $\ell$) and one seeks a surrogate loss $L$ with certain attributes, such as convexity, smoothness, etc.
In doing so, we may be interested in the minimum number of dimensions needed for such an indirect elicitation scheme; this is the notion of \emph{elicitation complexity}~\citep{lambert2008eliciting,fissler2016higher,frongillo2015elicitation}.
The two notions of elicitation complexity we use most are as follows: let $\eliccvx(\gamma)$ be the minimum dimension $d$ such that $\gamma$ can be indirectly elicited by a convex loss function $L: \R'^d \to \reals^\Y_+$, and $\elicpoly(\gamma)$ the minimum such $d$ where $L$ is polyhedral.
We will also discuss $\eliccts$, which is the complexity with respect to continuous, real-valued, non-locally-constant properties (\S~\ref{sec:1d}), and $\elicembed$, which is the minimum dimension for $\gamma$ to be embedded (Definition~\ref{def:embedding-dim}).

As a brief aside, we note that previous definitions of elicitation complexity have used a more straightforward link function, as the properties considered were single-valued~\citep{lambert2008eliciting,fissler2016higher,frongillo2015elicitation}.
In particular, if both $\gamma$ and $\Gamma$ are single-valued, these works simply require $\gamma = f \circ \Gamma$.
When $\Gamma$ is set-valued, one still straightforwardly requires $\gamma(p) = f(u)$ for all $u \in \Gamma(p)$.
The notion a calibrated link from Definition~\ref{def:links} is only required when both $\gamma$ and $\Gamma$ could be set-valued, as is the case in general for convex surrogates of finite properties.

\subsection{Important Examples}\label{sec:important-examples}
  Ranking and classification problems are well-known to the field of machine learning.
  The most studied classification problem is the elicitation of the mode.
  While the mode over $n$ outcomes is elicitable by 0-1 loss, it is known that eliciting the mode by a convex loss is much more difficult.
  In fact, the best convex surrogate that can be used to elicit the mode ``embeds'' the reports into $\reals^{n-1}$.
  Therefore, we say $\eliccvx($mode$) = n-1$.

@@

\section{Polyhedral Losses and Embeddings}
\label{sec:poly-loss-embed}

@@

\begin{definition}\label{def:embedding-dim}
  We say $\gamma$ is \emph{$d$-embeddable}, or just \emph{embeddable}, if such an $L$ exists and is convex.
  The \emph{embedding dimension} of a finite property $\gamma$, written $\elicembed(\gamma)$, is the minimum $d\geq 0$ such that $\gamma$ is $d$-embeddable.
\end{definition}

@@

\section{Embeddings via Real-Valued Losses}
\label{sec:1d}
